---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip) [![Travis-CI Build Status](https://travis-ci.com/thomascrines/statxplorer.svg?token=u9YrQxZw2wBxcEUArp4X&branch=master)](https://travis-ci.com/thomascrines/statxplorer) 

# statxplorer

Use statxplorer to download data from [Stat-Xplore](https://stat-xplore.dwp.gov.uk/webapi/jsf/login.xhtml). statxplorer can be used interactively, or as part of a [reproducible analytical pipeline](https://ukgovdatascience.github.io/rap_companion/).

*After development was started, we became aware of a similar package with the same name on [Github](https://github.com/olihawkins/statxplorer). As slightly different functionality is required in this package development is continuing, although the data-wrangling section of the first package was used as a basis for the statx_dataset function. We plan to change the name soon to avoid confusion.*

## Installation

Install opendatascot from GitHub with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("thomascrines/statxplorer")
```

If the above does not work, you can install from source:

1. Go to the statxplorer [repository](https://github.com/thomascrines/statxplorer) on GitHub
1. Click **Clone or download** then **Download ZIP**
1. Save the file locally (e.g. your H drive) and Unzip
1. Install with install.packages()

```{r source-installation, eval = FALSE}
install.packages("your/directory/statxplorer", repos = NULL,
                 type="source", lib = "your/R/package/library/directory")
```

Set-up
-----

### Api Key

Stat-Xplore API requires an API key, instructions on how to get one can be obtained in the [Stat-Xplore help](https://stat-xplore.dwp.gov.uk/webapi/online-help/Open-Data-API.html).

Once you have an API key, you can add it to an .Renviron file in your home directory by using statx_set_api_key:

```{r, eval=FALSE}
statxplorer::statx_set_api_key('YOUR_API_KEY')
```

You can also use statx_set_api_key to update with a new key.


*.Renviron only applies to RStudio. If you are not using RStudio, use 'Modify environment variables' or update .bash_profile in Mac to set environment variables.*

*You will also need to update the statx_request_bodies file from* 
```{r, eval=FALSE}
accessKey <- Sys.getenv("StatXploreApiKey")
```
*to refer to your variables.*

### Requests Folder

statxplorer is based on POST request bodies in JSON format, stored in files (more details below). To use statxplorer you must specify a folder path where request files will be stored, using statx_set_request_folder. Backslashes must be escaped. So for example to set the request folder to 'C:\\Documents\\StatXploreRequests' run `statx_set_request_folder('C:\\Documents\\StatXploreRequests')`

## Examples

### statx_add_request

statxplorer requires a JSON request body to send in a POST request to the Stat-Xplore API. To create one, either write one by hand by the Stat-Xplore guidelines, or use the webite to build a table and save it as Open Data API (.json). Once you have a .json file containing a request, run statx_add_request to add it to statexplorer, passing in your desired name for the file and the path of your existing file. For example, if you have a file with the path 'C:\\Downloads\\PensionCreditsNew001.json', that you want to save as 'PensionCredits', run `statx_add_request('PensionCredits', 'C:\\Downloads\\PensionCreditsNew001.json')`.

Alternatively, save and rename directly in the folder specified in statx_set_request_folder.

### statx_requests

To see a full list of the names of all available JSON requests, run `statx_requests()`.

### statx_print_request

To print the contents of a .json request file to the console pass the name of the request to statx_print_request. For example: `statx_print_request('PensionCredits')`.

### statx_delete_request

To delete a .json request file pass the name of the request to statx_delete_request. For example: `statx_delete_request('PensionCredits')`.

### statx_dataset

To return a complete dataset from Stat-Xplore based on an existing JSON request, pass the name of the request to statx_dataset. For example, `statx_dataset('PensionCredits')`.

### Editing requests

To edit a request, open directly for the request folder and edit in a text editor of your choice.
